Poincaré and SimBio: a versatile and extensible Python ecosystem for modeling systems.

### 2023-11-28 article.qmd
Expand reproducibility and performance

![Performance of the different softwares to solve a subset of BioModels models. (left) Run time for model BIOMD3 as a function of the number of output points. (right) Run time for different models ordered according to the geometric mean of the run times.](src/performance/figures/performance.png){#fig-runtime}

>>>>

![Performance of the different softwares to solve a subset of BioModels models. (left) Run time for model BIOMD3 as a function of the number of output points. (right) Run time for different models ordered according to the geometric mean of the run times, using 300 output points.](src/performance/figures/performance.png){#fig-runtime}

To test the reproducibility of results of SimBio,
we considered the first 100 curated models from BioModels.
As SimBio does not yet support the full SBML standard,
for instance, events,
we simulated only 60 of those.
Using the results from COPASI as ground truth,
SimBio reproduces the results within the solver tolerances.

To test the performance,
we simulated them with COPASI, Tellurium/RoadRunner and simbio.
For simbio, we considered both the NumPy and numba backends.
For COPASI and simbio, we used the LSODA solver,
while using the similar CVODE solver for tellurium.
In all cases, we used relative and absolute tolerances of  $10^-6$.
We measured separately three stages of the simulation:
the loading, first/cold run and subsequent/warm runs of each model.

For COPASI and Tellurium/RoadRunner,
we noted that its runtime depended on the number of evaluation points,
something that does not seem to happen with SimBio (@fig-runtime, left).
While SimBio's NumPy backend is slower than both COPASI and RoadRunner,
we obtained an order of magnitude speed-up using the numba backend putting it on par with them.
A user might have to consider the trade-off between compilation and run times,
as the compilation of the right-hand-side (RHS) code might take longer than the runtime itself,
and not be worth it for running the model only once.
Another speed-up in the runtime can be had
by switching the LSODA `scipy` solver for a more efficient `numbalsoda` implementation,
which avoids the Python interpreter between each of the integration steps.
This last combination beats all other methods,
which is also true for the other models we tested (@fig-runtime, right).
